hbm_paper_results

### 2024-04-19 paper_lbcc_results.Rmd
lbcc looking w/in feature

@@

```{r}
summary(lm(mean_cent ~ study, data=raw_centile_df.mean))
summary(aov(mean_cent ~ study, data=raw_centile_df.mean))
```
###Z-Scores

@@

```{r}
feat_to_test <- intersect(pheno_list, names(gamlss_centile_df.clean))

df_gamlss_feat_aov <- data.frame()
for(pheno in feat_to_test){
  # print(pheno)
  df <- tidy(aov(gamlss_centile_df.clean[[pheno]] ~ gamlss_centile_df.clean[["study"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov <- rbind(df_gamlss_feat_aov, df)
}

df_gamlss_feat_aov.study <- df_gamlss_feat_aov %>%
  group_by(pheno) %>%
  mutate(etasq = sumsq / sum(sumsq)) %>%
  ungroup() %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gamlss_feat_aov.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

@@

```{r}
gamlss.feat_to_test.z <- paste0(feat_to_test, ".z")
df_gamlss_feat_aov.z <- data.frame()
for(pheno in gamlss.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(gamlss_z_df.clean[[pheno]] ~ gamlss_z_df.clean[["study"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov.z <- rbind(df_gamlss_feat_aov.z, df)
}

@@

feat_aov.df <- full_join(df_gamlss_feat_aov.study, df_gam_feat_aov.study)
feat_aov.df <- full_join(feat_aov.df, df_raw_feat_aov.study)

@@

mean_cents.df <- full_join(gamlss_centile_df.mean, gam_centile_df.mean)
mean_cents.df <- full_join(mean_cents.df, raw_centile_df.mean)

@@

#Study-Level Mean Centiles ~ Study Ages

##Unharmonzied

```{r}
raw_centile_df.clean_long <- raw_centile_df.clean %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile")

#overall means
raw_centile_site.mean.df <- raw_centile_df.clean_long %>%
  group_by(study) %>%
  summarize(mean_raw_cent=mean(centile),
            median_raw_cent=median(centile),
            mean_age=mean(age_years),
            study_n= n()) %>%
  mutate(pheno_cat = "all")

summary(lm(mean_raw_cent ~ mean_age, data=raw_centile_site.mean.df))
summary(lm(mean_raw_cent ~ mean_age + study_n, data=raw_centile_site.mean.df))
```

## ComBatGAM

```{r}
gam_centile_df.clean_long <- gam_centile_df.clean %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile")

#overall means
gam_centile_site.mean.df <- gam_centile_df.clean_long %>%
  group_by(study) %>%
  summarize(mean_gam_cent=mean(centile),
            median_gam_cent=median(centile),
            mean_age=mean(age_years),
            med_age=median(age_years),
            study_n= n()) %>%
  mutate(pheno_cat = "all")

summary(lm(mean_gam_cent ~ mean_age, data=gam_centile_site.mean.df))
summary(lm(median_gam_cent ~ med_age, data=gam_centile_site.mean.df))
summary(lm(mean_gam_cent ~ mean_age + study_n, data=gam_centile_site.mean.df))
```

## ComBatLS

```{r}
gamlss_centile_df.clean_long <- gamlss_centile_df.clean %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile")

#overall means
gamlss_centile_site.mean.df <- gamlss_centile_df.clean_long %>%
  group_by(study) %>%
  summarize(mean_gamlss_cent=mean(centile),
            median_gamlss_cent=median(centile),
            mean_age=mean(age_years),
            med_age=median(age_years),
            study_n= n()) %>%
  mutate(pheno_cat = "all")

summary(lm(mean_gamlss_cent ~ mean_age, data=gamlss_centile_site.mean.df))
summary(lm(median_gamlss_cent ~ med_age, data=gamlss_centile_site.mean.df))
summary(lm(mean_gamlss_cent ~ mean_age + study_n, data=gamlss_centile_site.mean.df))
```

plots

```{r}
all_means.df <- full_join(gam_centile_site.mean.df, gamlss_centile_site.mean.df)
all_means.df <- full_join(all_means.df, raw_centile_site.mean.df)

all_means.df.long <- all_means.df %>%
  pivot_longer(cols=c("mean_gam_cent", "mean_gamlss_cent", "mean_raw_cent"), names_to="combat", values_to = "mean_cent") %>%
  mutate(combat=gsub("mean_(.+)_cent", "\\1", combat))

all_means.df.long %>%
  ggplot(aes(x=mean_age, y=mean_cent, color=combat)) +
  geom_point(aes(size=study_n)) +
  geom_smooth(method="lm")

all_means.df.long %>%
  ggplot(aes(x=med_age, y=mean_cent, color=combat)) +
  geom_point(aes(size=study_n), alpha=0.6) +
  geom_smooth(method="lm")
```

```{r}
all_medians.df.long <- all_means.df %>%
  pivot_longer(cols=c("median_gam_cent", "median_gamlss_cent", "median_raw_cent"), names_to="combat", values_to = "median_cent") %>%
  mutate(combat=gsub("median_(.+)_cent", "\\1", combat))

all_medians.df.long %>%
  ggplot(aes(x=med_age, y=median_cent, color=combat)) +
  geom_point(aes(size=study_n), alpha=0.6) +
  geom_smooth(method="lm")
```

@@

```{r}
study_info <- gam_centile_site.mean.df %>%
  dplyr::select(study, mean_age, med_age, study_n)

raw_centile_df.mean2 <- full_join(raw_centile_df.mean, study_info)
summary(lm(mean_cent ~ mean_age, data=raw_centile_df.mean2))

raw_centile_df.mean2 %>%
  ggplot(aes(x=mean_age, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

```{r}
gam_centile_df.mean2 <- full_join(gam_centile_df.mean, study_info)
summary(lm(mean_cent ~ mean_age, data=gam_centile_df.mean2))

gam_centile_df.mean2 %>%
  ggplot(aes(x=mean_age, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

```{r}
gamlss_centile_df.mean2 <- full_join(gamlss_centile_df.mean, study_info)
summary(lm(mean_cent ~ mean_age, data=gamlss_centile_df.mean2))

gamlss_centile_df.mean2 %>%
  ggplot(aes(x=mean_age, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

@@

all_subj_means.df <- full_join(gamlss_centile_df.mean2, gam_centile_df.mean2)
all_subj_means.df <- full_join(all_subj_means.df, raw_centile_df.mean2)

all_subj_means.df %>%
  ggplot(aes(x=mean_age, y=mean_cent, color=combat)) +
  geom_point(alpha=0.6) +
  geom_smooth(method="lm")

all_subj_means.df %>%
  ggplot(aes(x=mean_age, y=mean_cent, color=combat)) +
  geom_point(alpha=0.3) +
  geom_smooth(method="lm") +
  facet_wrap(~combat)

summary(lm(mean_cent ~ mean_age, data=raw_centile_df.mean2))
summary(lm(mean_cent ~ mean_age, data=gam_centile_df.mean2))
summary(lm(mean_cent ~ mean_age, data=gamlss_centile_df.mean2))
```

@@

##Age

delta ~ age

```{r}
# age
summary(lm(study_mean_abs.delta ~ study_mean_age_yrs, data=delta_sum.df_merged))
summary(lm(study_mean_delta ~ study_mean_age_yrs, data=delta_sum.df_merged))
```

>>>>

Deltas

```{r}
deltas_df <- fread(file=paste0(der_path, "/batch.study_pred_deltas.csv"))
deltas_df <- deltas_df %>% mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25)

dim(deltas_df)
min(deltas_df$age_years) == min(gamlss_centile_df$age_years)
```

Nature paper models to look at age effects in variance
from [https://github.com/brainchart/Lifespan/tree/main/Share](https://github.com/brainchart/Lifespan/tree/main/Share)

```{r}
paper_fit_sgmv <- readRDS("derivatives/nature_2020_mods/FIT_sGMV.rds")
paper_fit_gmv <- readRDS("derivatives/nature_2020_mods/FIT_GMV.rds")
paper_fit_wmv <- readRDS("derivatives/nature_2020_mods/FIT_WMV.rds")
paper_fit_ven <- readRDS("derivatives/nature_2020_mods/FIT_Ventricles.rds")
```

@@

```{r}
summary(lm(mean_cent ~ study, data=raw_centile_df.mean))
summary(aov(mean_cent ~ study, data=raw_centile_df.mean))
```

@@

```{r}
gamlss.feat_to_test <- intersect(pheno_list, names(gamlss_centile_df.clean))

df_gamlss_feat_aov <- data.frame()
for(pheno in gamlss.feat_to_test){
  # print(pheno)
  df <- tidy(aov(gamlss_centile_df.clean[[pheno]] ~ gamlss_centile_df.clean[["study"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov <- rbind(df_gamlss_feat_aov, df)
}

df_gamlss_feat_aov.study <- df_gamlss_feat_aov %>%
  group_by(pheno) %>%
  mutate(etasq = sumsq / sum(sumsq)) %>%
  ungroup() %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gamlss_feat_aov.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

@@

```{r}
gamlss.feat_to_test.z <- paste0(gamlss.feat_to_test, ".z")
df_gamlss_feat_aov.z <- data.frame()
for(pheno in gamlss.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(gamlss_z_df.clean[[pheno]] ~ gamlss_z_df.clean[["study"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov.z <- rbind(df_gamlss_feat_aov.z, df)
}

@@

#put all data frames into list
df_list.tmp <- list(df_gamlss_feat_aov.study, df_gam_feat_aov.study, df_raw_feat_aov.study)

#merge all data frames in list
feat_aov.df <- df_list.tmp %>% reduce(full_join)

@@

#put all data frames into list
df_list.tmp <- list(gamlss_centile_df.mean, gam_centile_df.mean, raw_centile_df.mean)

#merge all data frames in list
mean_cents.df <- df_list.tmp %>% reduce(full_join)

@@

#Feature Centile ~ Mean Study Age

Get study info
```{r}
study_info <- raw_centile_df.clean %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
  group_by(study) %>%
  summarize(study_mean_age=mean(age_years),
            study_median_age=median(age_years),
            study_n= n()) %>%
  dplyr::select(study, study_mean_age, study_median_age, study_n)
```

W/in feat lms

##Unharmonized

```{r}
raw_centile_df.clean_study <- full_join(raw_centile_df.clean, study_info)

df_raw_feat_lm <- data.frame()
for(pheno in raw.feat_to_test){
  # print(pheno)
  df <- tidy(aov(raw_centile_df.clean_study[[pheno]] ~ raw_centile_df.clean_study[["study_mean_age"]]))
  df$pheno <- pheno
  df_raw_feat_lm <- rbind(df_raw_feat_lm, df)
}

df_raw_feat_lm <- df_raw_feat_lm %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(raw.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_raw_feat_lm %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

df_raw_feat_lm %>%
  ggplot() +
  geom_histogram(aes(x=statistic))
```

##ComBat-GAM

```{r}
gam_centile_df.clean_study <- full_join(gam_centile_df.clean, study_info)

df_gam_feat_lm <- data.frame()
for(pheno in gam.feat_to_test){
  # print(pheno)
  df <- tidy(aov(gam_centile_df.clean_study[[pheno]] ~ gam_centile_df.clean_study[["study_mean_age"]]))
  df$pheno <- pheno
  df_gam_feat_lm <- rbind(df_gam_feat_lm, df)
}

df_gam_feat_lm <- df_gam_feat_lm %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gam_feat_lm %>%
  ggplot() +
  geom_histogram(aes(x=statistic))
```

```{r}
gamlss_centile_df.clean_study <- full_join(gamlss_centile_df.clean, study_info)

df_gamlss_feat_lm <- data.frame()
for(pheno in gamlss.feat_to_test){
  # print(pheno)
  df <- tidy(aov(gamlss_centile_df.clean_study[[pheno]] ~ gamlss_centile_df.clean_study[["study_mean_age"]]))
  df$pheno <- pheno
  df_gamlss_feat_lm <- rbind(df_gamlss_feat_lm, df)
}

df_gamlss_feat_lm <- df_gamlss_feat_lm %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gamlss_feat_lm %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

df_gamlss_feat_lm %>%
  ggplot() +
  geom_histogram(aes(x=statistic))
```

*statistic ~ effect size?

```{r}
df_raw_feat_lm$combat <- "raw"
df_gam_feat_lm$combat <- "gam"
df_gamlss_feat_lm$combat <- "gamlss"

#put all data frames into list
df_list.tmp <- list(df_raw_feat_lm, df_gam_feat_lm, df_gamlss_feat_lm)

#merge all data frames in list
all_feat_lm <- df_list.tmp %>% reduce(full_join)

all_feat_lm %>%
  ggplot() +
  geom_histogram(aes(x=statistic, fill=combat, color=combat), position="identity", alpha=0.5, bins=50)
```

@@

```{r}
raw_centile_df.mean2 <- full_join(raw_centile_df.mean, study_info)
summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))

raw_centile_df.mean2 %>%
  ggplot(aes(x=study_mean_age, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

```{r}
gam_centile_df.mean2 <- full_join(gam_centile_df.mean, study_info)
summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))

gam_centile_df.mean2 %>%
  ggplot(aes(x=study_mean_age, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")

gam_centile_df.mean2 <- gam_centile_df.mean2 %>%
  mutate(age_diff = (age_days/365.25) - study_mean_age)

summary(lm(mean_cent ~ abs(age_diff), data=gam_centile_df.mean2))
gam_centile_df.mean2 %>%
  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

```{r}
gamlss_centile_df.mean2 <- full_join(gamlss_centile_df.mean, study_info)
summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))

gamlss_centile_df.mean2 %>%
  ggplot(aes(x=study_mean_age, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")

gamlss_centile_df.mean2 <- gamlss_centile_df.mean2 %>%
  mutate(age_diff = (age_days/365.25) - study_mean_age)

summary(lm(mean_cent ~ abs(age_diff), data=gamlss_centile_df.mean2))
gamlss_centile_df.mean2 %>%
  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

@@

#put all data frames into list
df_list.tmp <- list(gamlss_centile_df.mean2, gam_centile_df.mean2, raw_centile_df.mean2)

#merge all data frames in list
all_subj_means.df <- df_list.tmp %>% reduce(full_join)

all_subj_means.df %>%
  ggplot(aes(x=study_mean_age, y=mean_cent, color=combat)) +
  geom_point(alpha=0.6) +
  geom_smooth(method="lm")

all_subj_means.df %>%
  ggplot(aes(x=study_mean_age, y=mean_cent, color=combat)) +
  geom_point(alpha=0.3) +
  geom_smooth(method="lm") +
  facet_wrap(~combat)

summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))
summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))
summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))
```

@@

#ComBat-GAM vs ComBatLS: Delta Centile

clean delta df

```{r}
cent_deltas_df <- deltas_df %>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles

#get subject-level means
mean_cent_deltas_df <- cent_deltas_df %>%
  pivot_longer(cols=any_of(paste0("delta.", pheno_list)), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_delta_cent=(mean(centile))*100, #convert units to centiles space
            abs_mean_delta_cent=(mean(abs(centile)))*100)

mean_cent_deltas_df_study <- full_join(mean_cent_deltas_df, study_info)
```

check mean subj delta ~ study mean age

```{r}
delta_lm <- lm(mean_delta_cent ~ study_mean_age + study_n, data=mean_cent_deltas_df_study)
summary(delta_lm)

mean_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=mean_delta_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(delta_lm)[["study_mean_age"]], 
              intercept = coef(delta_lm)[["(Intercept)"]])+
  theme(legend.position="none")
```

mean abs delta ~ study mean age

```{r}
abs_lm <- lm(abs_mean_delta_cent ~ study_mean_age + study_n, data=mean_cent_deltas_df_study)
summary(abs_lm)

mean_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_mean_delta_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_lm)[["study_mean_age"]], 
              intercept = coef(abs_lm)[["(Intercept)"]])+
  theme(legend.position="none")

mean_cent_deltas_df_study <- mean_cent_deltas_df_study %>%
  mutate(age_diff = (age_days/365.25) - study_mean_age)

abs_lm2 <- lm(abs_mean_delta_cent ~ study_mean_age + study_n + abs(age_diff), data=mean_cent_deltas_df_study)
summary(abs_lm2)

mean_cent_deltas_df_study %>%
  ggplot(aes(x=abs(age_diff), y=abs_mean_delta_cent)) +
  geom_point(alpha=0.05) +
  geom_smooth(method="lm")
```

*options: rerun from the beginning w/o UKB? Or just control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting?

###Global Features
Look at global features

```{r}
glob_cent_deltas_df <- cent_deltas_df %>%
  dplyr::select(study, age_years, sex, participant, fs_version, paste0("delta.", vol_list_global)) %>%
  mutate(across(paste0("delta.", vol_list_global), abs, .names = "abs_{.col}"))

glob_cent_deltas_df_long <- glob_cent_deltas_df %>%
  pivot_longer(cols=starts_with("abs"), names_to="pheno", values_to = "abs_delta") %>%
  mutate(pheno=gsub("abs_delta.", "", pheno))

glob_cent_deltas_df_long %>%
  ggplot() +
  geom_boxplot(aes(y=abs_delta, color=pheno), outlier.shape = NA) + 
scale_y_continuous(limits = quantile(glob_cent_deltas_df_long$abs_delta, c(0.1, 0.9)))
```
```{r}
glob_cent_deltas_df_study <- full_join(glob_cent_deltas_df, study_info)

abs_gmv_lm <- lm(abs_delta.GMV ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)
summary(abs_gmv_lm)

glob_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_delta.GMV)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_gmv_lm)[["study_mean_age"]], 
              intercept = coef(abs_gmv_lm)[["(Intercept)"]])+
  theme(legend.position="none") +
  scale_y_continuous(limits = c(0, 0.1))

abs_sgmv_lm <- lm(abs_delta.sGMV ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)
summary(abs_sgmv_lm)
glob_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_delta.sGMV)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_sgmv_lm)[["study_mean_age"]], 
              intercept = coef(abs_sgmv_lm)[["(Intercept)"]])+
  theme(legend.position="none")+
  scale_y_continuous(limits = c(0, 0.1))

abs_wmv_lm <- lm(abs_delta.WMV ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)

summary(gam(abs_delta.WMV ~ pb(study_mean_age) + study_n, data=glob_cent_deltas_df_study))

summary(abs_wmv_lm)
glob_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_delta.WMV)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_wmv_lm)[["study_mean_age"]], 
              intercept = coef(abs_wmv_lm)[["(Intercept)"]])+
  theme(legend.position="none")+
  scale_y_continuous(limits = c(0, 0.1))

abs_ven_lm <- lm(abs_delta.Ventricles ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)
summary(abs_ven_lm)
glob_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_delta.Ventricles)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_ven_lm)[["study_mean_age"]], 
              intercept = coef(abs_ven_lm)[["(Intercept)"]])+
  theme(legend.position="none")+
  scale_y_continuous(limits = c(0, 0.1))

glob_cent_deltas_df_long <- full_join(glob_cent_deltas_df_long, study_info)

glob_cent_deltas_df_long <- glob_cent_deltas_df_long %>%
  mutate(age_diff=age_years-study_mean_age)

glob_cent_deltas_df_long %>%
  ggplot(aes(x=study_mean_age, y=abs_delta, color=pheno)) +
  geom_point(alpha=0.4) +
  geom_smooth(method="lm") +
  facet_wrap(~pheno, scales="free_y")
```

####Compare w/ Nature Paper
Compare to Nature paper age effects in sigma

```{r}
#gmv
paper_fit_gmv$param$sigma$fixef
gmv_b1 <- unname(paper_fit_gmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 3))1"])
gmv_b2 <- unname(paper_fit_gmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 3))2"])

#formula
gmv_age_sig <- parse(text = paste0(gmv_b1, "*x^(-2) + ", gmv_b2,"*x^(3)"))
gmv_age_sig_der <- D(gmv_age_sig,"x")

#sgmv
paper_fit_sgmv$param$sigma$fixef
sgmv_b1 <- unname(paper_fit_sgmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2))1"])
sgmv_b2 <- unname(paper_fit_sgmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2))2"])

sgmv_age_sig <- parse(text = paste0(sgmv_b1, "*x^(-2) + ", sgmv_b2,"*x^(-2)*log(x)"))
sgmv_age_sig_der <- D(sgmv_age_sig,"x")

#wmv
paper_fit_wmv$param$sigma$fixef
wmv_b1 <- unname(paper_fit_wmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 2))1"])
wmv_b2 <- unname(paper_fit_wmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 2))2"])

wmv_age_sig <- parse(text = paste0(wmv_b1, "*x^(-2) + ", wmv_b2,"*x^(s)"))
wmv_age_sig_der <- D(wmv_age_sig,"x")

#ven
paper_fit_ven$param$sigma$fixef
ven_b1 <- unname(paper_fit_ven$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2, -2))1"])
ven_b2 <- unname(paper_fit_ven$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2, -2))2"])
ven_b3 <- unname(paper_fit_ven$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2, -2))3"])

ven_age_sig <- parse(text = paste0(ven_b1, "*x^(-2) + ", ven_b2,"*x^(-2)*log(x) + ", ven_b3, "*x^(-2)*log(x)^(2)"))
ven_age_sig_der <- D(ven_age_sig,"x")
```
try summing abs(betas) within each pheno to get overall magnitude of age effects

```{r}
paper_sigma_age <- data.frame("pheno" = c("GMV", "sGMV", "WMV", "Ventricles"),
                              "paper_abs_betas_sum" = c(sum(abs(gmv_b1), abs(gmv_b2)), 
                                                  sum(abs(sgmv_b1), abs(sgmv_b2)),
                                                  sum(abs(wmv_b1), abs(wmv_b2)),
                                                  sum(abs(ven_b1), abs(ven_b2), abs(ven_b3))),
                              "paper_betas_sum" = c(sum(gmv_b1, gmv_b2), 
                                                  sum(sgmv_b1, sgmv_b2),
                                                  sum(wmv_b1, wmv_b2),
                                                  sum(ven_b1, ven_b2, ven_b3)),
                              "lm_beta" = c(coef(abs_gmv_lm)[["study_mean_age"]],
                                            coef(abs_sgmv_lm)[["study_mean_age"]],
                                            coef(abs_wmv_lm)[["study_mean_age"]],
                                            coef(abs_ven_lm)[["study_mean_age"]]),
                              "quick_var_range" = c((38.85552-16.93754), (0.2380905-0.1447285), (29.34474-13.4674), (17.63019-0.3520969))
                              )
paper_sigma_age %>%
  ggplot() +
  geom_col(aes(x=pheno, y=paper_abs_betas_sum, fill=pheno))

paper_sigma_age %>%
  ggplot() +
  geom_col(aes(x=pheno, y=paper_betas_sum, fill=pheno))

paper_sigma_age %>%
  ggplot() +
  geom_point(aes(x=abs(lm_beta), y=paper_betas_sum, color=pheno))

paper_sigma_age %>%
  ggplot() +
  geom_point(aes(x=abs(lm_beta), y=quick_var_range, color=pheno))
```

###W/In Cortical Features

Looking at averages w/in each cortical feature since they have different trajectories of variance across the lifespan (see Nature paper)

####Regional Volumes

```{r}
#get subject-level means
mean_vol_cent_deltas_df <- cent_deltas_df %>%
  pivot_longer(cols=any_of(paste0("delta.", vol_list_regions)), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_delta_vol_cent=(mean(centile))*100, #convert units to centiles space
            abs_mean_delta_vol_cent=(mean(abs(centile)))*100) %>%
  dplyr::select(!(starts_with("delta")))

mean_vol_cent_deltas_df_study <- full_join(mean_vol_cent_deltas_df, study_info)

abs_vol_lm <- lm(abs_mean_delta_vol_cent ~ study_mean_age + study_n, data=mean_vol_cent_deltas_df_study)
summary(abs_vol_lm)

mean_vol_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_mean_delta_vol_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_vol_lm)[["study_mean_age"]], 
              intercept = coef(abs_vol_lm)[["(Intercept)"]])+
  theme(legend.position="none")
```

####Regional Thickness

```{r}
#get subject-level means
mean_ct_cent_deltas_df <- cent_deltas_df %>%
  pivot_longer(cols=any_of(paste0("delta.", ct_list)), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_delta_ct_cent=(mean(centile))*100, #convert units to centiles space
            abs_mean_delta_ct_cent=(mean(abs(centile)))*100) %>%
  dplyr::select(!(starts_with("delta")))

mean_ct_cent_deltas_df_study <- full_join(mean_ct_cent_deltas_df, study_info)

abs_ct_lm <- lm(abs_mean_delta_ct_cent ~ study_mean_age + study_n, data=mean_ct_cent_deltas_df_study)
summary(abs_ct_lm)

mean_ct_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_mean_delta_ct_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_ct_lm)[["study_mean_age"]], 
              intercept = coef(abs_ct_lm)[["(Intercept)"]])+
  theme(legend.position="none")
```

much bigger than anticipated based on Nature paper!!? is combatLS overfitting?

```{r}
#get subject-level means
mean_ct_cent_gamlss_df <- gamlss_centile_df.clean %>%
  pivot_longer(cols=any_of(ct_list), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_ct_cent=(mean(centile))*100)
```

```{r}
mean_ct_cent_gamlss_df <- full_join(mean_ct_cent_gamlss_df, study_info)
summary(lm(mean_ct_cent ~ study_mean_age, data=mean_ct_cent_gamlss_df))
summary(lm(mean_ct_cent ~ study_mean_age + study_n, data=mean_ct_cent_gamlss_df))

# mean_ct_cent_gamlss_df %>%
#   ggplot(aes(x=study_mean_age, y=mean_ct_cent)) +
#   geom_point(aes(color=study), alpha=0.2) +
#   geom_smooth(method = "lm") +
#   theme(legend.position="none")

#get subject-level means
mean_ct_cent_gam_df <- gam_centile_df.clean %>%
  pivot_longer(cols=any_of(ct_list), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_ct_cent=(mean(centile))*100)

mean_ct_cent_gam_df <- full_join(mean_ct_cent_gam_df, study_info)
summary(lm(mean_ct_cent ~ study_mean_age, data=mean_ct_cent_gam_df))
summary(lm(mean_ct_cent ~ study_mean_age + study_n, data=mean_ct_cent_gam_df))

# mean_ct_cent_gam_df %>%
#   ggplot(aes(x=study_mean_age, y=mean_ct_cent)) +
#   geom_point(aes(color=study), alpha=0.2) +
#   geom_smooth(method = "lm") +
#   theme(legend.position="none")
```

####Regional SA

```{r}
#get subject-level means
mean_sa_cent_deltas_df <- cent_deltas_df %>%
  pivot_longer(cols=any_of(paste0("delta.", sa_list)), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_delta_sa_cent=(mean(centile))*100, #convert units to centiles space
            abs_mean_delta_sa_cent=(mean(abs(centile)))*100) %>%
  dplyr::select(!(starts_with("delta")))

mean_sa_cent_deltas_df_study <- full_join(mean_sa_cent_deltas_df, study_info)

abs_sa_lm <- lm(abs_mean_delta_sa_cent ~ study_mean_age + study_n, data=mean_sa_cent_deltas_df_study)
summary(abs_sa_lm)

mean_sa_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_mean_delta_sa_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_sa_lm)[["study_mean_age"]], 
              intercept = coef(abs_sa_lm)[["(Intercept)"]])+
  theme(legend.position="none")
```

####Compare

```{r}
df_list.tmp <- list(mean_sa_cent_deltas_df_study, mean_ct_cent_deltas_df_study, mean_vol_cent_deltas_df_study)
all_reg_deltas <- df_list.tmp %>% reduce(full_join)
```

```{r}
all_reg_deltas_long <- all_reg_deltas %>%
  pivot_longer(cols=starts_with("abs_mean_delta"), names_to="pheno_type", values_to = "abs_cent_delta") %>%
  mutate(pheno_type=gsub("abs_mean_delta_|_cent", "", pheno_type))
